Our Future Nuclear Data Needs

### 2019-01-01 nuclear_data_needs/drafts/nuclear_data_needs-121618-v1.tex
Adding old draft versions of review article

\newcommand{\keff}{k$_{\textrm{eff}}$}

@@

%Authors, affiliations address.
\author{Lee A. Bernstein$^{1,2}$, David A. Brown$^3$, \\Arjan J. Koning$^4$, Bradley T. Rearden$^5$,  \\Catherine E. Romano$^5$, Alejandro A. Sonzogni$^3$, Andrew S. Voyles$^{2}$, and Walid  Younes$^6$
\affil{$^1$Nuclear Science Division, Lawrence Berkeley National Laboratory,  Berkeley, CA 94720, USA; email: labernstein@lbl.gov}
\affil{$^2$Department of Nuclear Engineering, University of California, Berkeley,  Berkeley, CA 94720, USA}
\affil{$^3$National Nuclear Data Center, Brookhaven National Laboratory, Upton, NY 11973, USA}
\affil{$^4$Nuclear Data Section, International Atomic Energy Agency, 1400 Vienna, Austria}
\affil{$^5$Reactor and Nuclear Systems Division, Oak Ridge National Laboratory, Oak Ridge, TN 37830, USA}
\affil{$^6$Lawrence Livermore National Laboratory L-414, Livermore, CA 94551, USA}}

%Abstract
\begin{abstract}
A well-established knowledge of nuclear phenomena including fission, reaction cross sections, and structure/decay properties, is critical for applications ranging from the design of new reactors, to non-proliferation, to the production of radioisotopes for the diagnosis and treatment of illness.
However, the lack of a well-quantified, predictive theoretical capability means that most nuclear observables must be measured directly and used to \enquote{calibrate} empirical models, which in turn, provide the data needed for these applications.
 In many cases there is either a lack of the data needed to guide the models, or the results of the different measurements are discrepant.
 This has led to the development of evaluation methodologies to provide recommended values and uncertainties.
 In this article  the nuclear data evaluation process and the international community that carries it out are presented.
 We then discuss new measurements and improved theory and/or modeling needed to address future challenges in applied nuclear science. 
\end{abstract}

@@

Low-energy nuclear science (LENS) \enquote{straddles the fence} between curiosity- and application-driven pursuits.
 On the curiosity-driven side, low-energy nuclear structure offers a unique laboratory to examine the interplay between single-particle and collective behavior and to explore the transition from quantum to continuum behavior in a mesoscopic setting.
LENS also plays an important role in nuclear astrophysics, improving models of stellar energy generation and allowing isotopic abundances to be used to inform models of the astrophysical environments where heavy nuclei are formed.
 On the applied side, a well-quantified knowledge of low-lying nuclear structure and reactions is needed to model energy generation and isotope production for medical and industrial uses and for the national security and counter-proliferation communities.

One hallmark of LENS is that most theoretical descriptions are descriptive rather than rigorously predictive at the level of accuracy and over the entire range of energies etc. needed for real world applications.
 As a result, targeted measurements coupled to a data evaluation process are used to produce databases of recommended values that can be employed by the user communities.
 The evaluation processes range from close adherence to experimental results in the case of low energy nuclear structure and decay data, to experiment-guided modeling in the case of nuclear reactions.
This work is carried out by subject matter experts who usually come from the field utilizing the data.  These individuals spend many years studying under experienced evaluators in order to develop the highly-specialized skills required for the evaluation process.

This manuscript starts with a description of the nuclear data evaluation process, including the domestic and international organizations that help organize evaluation efforts.  It  then describes ways to improve the modeling of (n,x) reactions on nuclides important for applications.  This includes providing a better understanding of fission, improving the modeling of inelastic neutron scattering and better determining the properties of the highly-excited continuum states which determine which reactions dominate at a given excitation energy.
 Following this is a discussion of needs for specific applications including national security/non-proliferation, medical isotope production and nuclear energy.
 It ends by introducing a recently-formed Nuclear Data Interagency Working Group that's formulating a national plan to address high-priority needs, and by presenting ideas about how to change the way that we educate evaluators to ensure that a well-trained workforce is available to meet these important needs.

\subsection{The Nuclear Data Pipeline}
\label{sec:pipeline}

Before diving into the data needs, it is instructive to understand the \enquote{Nuclear Data Pipeline} that takes experimental or theoretical results and prepares them for use in applications.
 This process is illustrated in \autoref{fig:pipeline} and is roughly grouped into four steps: 1) compilation, 2) evaluation, 3) processing, and 4) validation.
 This process ends with the users and their application, but actually begins with the performance of carefully-designed measurements with well-defined uncertainties.
 Given this, those funding or performing the original experiment would be wise to understand the entire nuclear data workflow in order to ensure that new results find their way to the intended user.
 For example, if a new result forces changes to data formats (step 2) or processing (step 3) or to application codes (step 4 \& eventual user), this fact should be considered when developing a new activity.

@@

The nuclear structure compilation database is called the Experimental Unevaluated Nuclear Data List (XUNDL) \cite{XUN}.
  The nuclear data in XUNDL is organized by nuclide, meaning that a single publication could lead to the production of more than one XUNDL dataset.
 Each XUNDL dataset is treated as a \enquote{stand-alone} work and is not required to agree with existing, evaluated nuclear data for the nuclide it concerns.
 However, during XUNDL compilation basic consistency checks are performed and any internal inconsistencies are pointed out to the author to allow them be reconciled in the database.
 XUNDL is used by both nuclear structure evaluators to aid in their work, and by nuclear structure researchers as a quick way to get access to data from current publications in order to guide their own research.

@@

The next step in the process is evaluation itself.
 In the case of nuclear structure data this involves reconciling multiple data set types (e.g., prompt gamma and particle spectra, decay data etc.) and deriving a recommended set of adopted values.
 This data forms the basis of the Evaluated Nuclear Structure Data File (ENSDF) \cite{tepel1984ensdf}.
 While most of the ENSDF evaluation is performed in the US, there is a significant and growing component coming from evaluators throughout the world.
 The governing body that determines the rules for the evaluation process is the Nuclear Structure and Decay Data Network administered by the International Atomic Energy Agency (IAEA-NSDD).

In the case of nuclear reaction data, the process is markedly different.
 The data from EXFOR is used to guide a physics-based model calculation through a variety of variance minimization procedures, resulting in best estimates of mean values and their uncertainties, including covariances, which are put into the one or more regional libraries such as the recently release of the Evaluated Nuclear Data File (ENDF), ENDF/B-VIII.0 \cite{endf8_0}.
 Most of the effort put into the production of ENDF is provided by applications-oriented users due to its importance to national security, international counter-proliferation and nuclear energy.
 A consequence of this application focus is that the vast majority of the data in ENDF concerns neutron-induced reactions at either thermal (25 meV), fast fission or 14.1 MeV.
 This has profound implications for isotope production since many of the nuclides are produced through charged-particle induced reactions.

While ENSDF and ENDF contain a vast quantity of nuclear data, there are significant amounts of data that are either not contained in them because they \enquote{fall between} nuclear structure and nuclear reactions, or they are present in a format that is not well-suited to their use in applications.
 Examples of this include the recently updated Atlas of Neutron Resonances \cite{Mug18} which contains neutron capture resonance widths, centroids and cross sections and the Evaluated Gamma Activation File of (EGAF) capture gamma-rays \cite{Fir15} and the Atlas of Inelastic Scattering of Reactor Fast Neutrons \cite{Hur18}.
 Another very important nuclear data resource for applications is the Reference Input Parameter Library (RIPL) \cite{Capote2009} which contains both discrete (distilled from ENSDF) and continuous (Nuclear Level Densities and Gamma-strength function) nuclear structure data needed for accurate reaction modeling.

Evaluations in hand, the data has to be processed into a form suitable for use in a particular application.
 This is a surprisingly non-trivial part of the process, requiring a deep knowledge of the physics used in the evaluation process, the formatting limitations of intermediate data formats, and the capabilities of downstream codes that will use the nuclear data.
 Processing is also needed for the validation of nuclear reaction data using the integral benchmarks described below.
 Another role of processing is in the creation of application-specific databases, such as the Medical Internal Radiation Dose (MIRD) \cite{MIRD} database used to guide medical treatment and diagnosis.

The final step in the Pipeline is validation.
 In this step, user codes and user-defined problems are used to validate that the data performs as expected.
 The class of user problems most relied on are so-called \enquote{integral benchmark} data that depend on many different types of data including cross sections and emitted particle spectra and are also extremely precisely known.
 The canonical example of such a benchmark is a critical assembly, where the ratio of neutron production to neutron loss (\keff) is known to several significant figures.
 However, there are many other kinds of integral benchmarks used to test neutron interaction data alone, including reaction rate measurements and 14 MeV and fission spectrum transmission experiments, and many more for other forms of nuclear data.

@@

Much of the work of nuclear data evaluation takes place as a part of international collaborations covered under the Organization of Economic Cooperation and Development's Nuclear Energy Agency (NEA) and the International Atomic Energy Agency (IAEA).
 The Nuclear Energy Agency (NEA) addresses the nuclear technology interests of its 33 member states and, in the area of nuclear reaction data, the Working Party on Evaluation Coordination (WPEC) is the main forum for collaborative effort between the nuclear data library projects from the NEA countries, namely ENDF/B (United States), JENDL (Japan), JEFF (NEA), TENDL (Europe), BROND (Russia) as well as the non-OECD file project CENDL (China).

A powerful instrument of WPEC to drive progress in nuclear data is the so-called Subgroup (SG): members of WPEC identify a common area in nuclear data that requires improvement, and if enough support from the various data library projects is present, a Subgroup is formed.
 Subgroups typically operate on a 3-5 year time frame.
A recent successful example of a WPEC subgroup, with specific relevance to the US, includes the large-scale horizontal CIELO effort (SG40) on worldwide nuclear data evaluation for the most important fission energy related materials.
The CIELO initiative has led to new ENDF/B-VIII.0 evaluations for \ce{^{235,238}U}, \ce{^{56}Fe} and \ce{^{16}O} among others.  Other successful subgroups include one centered on developing a new Generalized Nuclear Database Structure (GNDS) format (SG38) \cite{SG38}, and one focused  on covariance adjustment for improvement of nuclear data files (SG39) \cite{SG39}.

WPEC also hosts long-term Expert Groups.  Current groups are working on the development of the Generalized Nuclear Database Structure (GNDS) format mentioned above that will provide a data library interface between nuclear physics and applications more modern than the ENDF-6 format, and the High-Priority Request List which assembles the most important nuclear data requests from applications in a unified format, to stimulate experimentalists and evaluators to provide these data.  A full list of past and current subgroups of WPEC is available at Ref. \cite{WPEC}.

A key player in nuclear data evaluation is the the International Atomic Energy Agency (IAEA), which covers the interests of its 170 member states.
The main task of the IAEA Nuclear Data Section is to provide fundamental nuclear databases for basic and applied use, with data originating from experiments and theoretical simulations covering both nuclear structure and nuclear reaction data.
 An important collaboration coordinated by the IAEA is the Nuclear Reaction Data Center Network, which is responsible for keeping the EXFOR database of experimental nuclear reaction data up to date. The NNDC is responsible for the US input to EXFOR.

In addition, the IAEA organizes Coordinated Research Projects (CRP) and technical meetings as instruments to align international nuclear data efforts towards the production of validated databases ready for applied use.
 Examples of recent and current CRP's include a 2018 venture centered on nuclear data for primary radiation damage, which has been completed, an ongoing effort to improve nuclear model parameters for fission reaction calculations by modern nuclear model codes such as EMPIRE \cite{Herman2007}, CCONE \cite{Iwa16}, COH3 \cite{KAWANO2010} and TALYS \cite{Koning2012}, and an effort to create a first-ever evaluated database of radiative strength.
 In 2019, a CRP on fission yields will start which aims to produce updated fission yield libraries for the major actinides to respond to requests from reactor technology, safeguards and non-proliferation.

@@

For fission theory, we can make a distinction in practice between description of the fission process up to scission (pre-scission physics), and the events that follow scission (post-scission physics).
Eventually, one could envision a comprehensive model that accurately follows the entire fission process from the formation of the parent nucleus to the last decays of the final product nuclei within a single internally consistent theoretical framework.
Such a model is not yet feasible, so different physics models are used to treat pre- and post-scission processes.
A wide range of theoretical approaches have been developed since the late 1930's to describe pre-scission physics, including scission-point models \cite{Wil76,Lem15}, microscopic-macroscopic models \cite{Mol15} with random-walk dynamics \cite{War17}, Langevin dynamics \cite{Sie17}, Time-dependent Hartree-Fock and Hartree-Fock-Bogoliubov theory \cite{Koo76,Neg78,Tan15,God16}, the time-dependent superfluid local density approximation \cite{Ste11,Bul13,Ste15,Bul16}, and the time-dependent generator-coordinate method \cite{Ber91,Gou05,Rei16}.Several models have also been developed to describe post-scission physics and built into computer codes, such as GEF \cite{Schm16}, FIFRELIN \cite{Lit15}, CGMF \cite{Tal14}, and FREYA \cite{Ran09,Ver18}.

The division between pre- and post-scission physics in modeling efforts suggests a corresponding division in nuclear data needs to improve our understanding of the fission process.
Post-scission observables are generally easier to measure, and include the properties of fission fragments and the neutrons and gamma rays they emit.
 Recently, Bertsch et al. \cite{Ber15} have suggested a subset of fission data that tend to have relatively small uncertainties and could therefore be used to test and validate fission models.
These data consist of fission barriers, fragment mass distributions, fragment total kinetic energies, spontaneous-fission lifetimes, and fission isomer excitation energies.
In addition to these quantities, the angular distribution of fission fragments can also provide useful information about the angular momentum of the fissioning nucleus and further constrain model calculations.

@@

Both pre-scission and post-scission theory and data will be needed to reach a satisfactory understanding of the fission process, and thereby make reliable predictions of nuclear quantities that cannot be measured in the laboratory.
Pre-scission models can provide input to post-scission models wherever the data are lacking, and post-scission models can be used to improve pre-scission calculations using a wealth of available experimental data.

The ultimate goal for theorists, a comprehensive description of fission starting from protons, neutrons, and their interactions that provides accurate values for all fission data of interest, remains elusive.  Many questions about the fission process remain open, such as: how the initial partition of energy between kinetic and internal excitation occurs, what the relevant collective and single-particle degrees of freedom are near scission, and how they evolve and couple throughout the process? Lastly, a compelling questions that remains is how to best quantify uncertainties (both epistemic and aleatory) in fission calculations.  Improved experiments and theoretical models are needed to address all of these questions, and to provide the deeper understanding of the fission process needed for most applications.

@@

\begin{figure}[b]
 \centering
%  \includegraphics[clip=true,trim=1.5in 3.4in 1.8in 3.5in, scale=0.8]{./figures/variation_curve.pdf}
 \includegraphics[width=0.7\linewidth]{rotated_swings.pdf}

@@

Elastic and inelastic scattering have a profound effect on neutron populations in any nuclear application by changing the directions and energies of the neutrons in the system.
 However, improvements in neutron scattering data have been hampered by the fact that they do not produce easily measured signals like fission, or a radiochemical observable like neutron capture.
 This has resulted in significant disagreements between different reaction evaluations.
 These disagreements manifest themselves as compensating errors that are resistant to the validation process.
 This issue has been recognized by the nuclear data community since the early days of the WPEC \cite{Row89} and more recently by Maslov \cite{Mas11}.

One compelling example of compensating errors involves the Jezebel \ce{^{239}Pu} critical assembly (PMF1).
 Models of Jezebel using LANL CIELO-1 (ENDF/B-VIII.0) and CEA CIELO-2 (JEFF-3.3) reproduce a unity value of $k_{eff}$ to better than 1 part in 104.
 However, when the authors substituted one nuclear quantity at a time between the two databases, the value of $k_{eff}$ varied considerably.
 The greatest changes came from changing the elastic and inelastic scattering cross sections, which led to changes in the $k_{eff}$ of -500 pcm and +704 pcm, respectively.  This is shown in \autoref{fig:rotated_swings}.  Similar examples of such compensating errors were also seen in earlier work by Bauge \cite{Bau12}.

The Collaborative International Evaluated Library Organization (CIELO) Coordinated Research Program referred to above \cite{Cha18a} was started to have the international nuclear data evaluation community agree on a set of evaluated cross sections and output spectra for six critically-important nuclei: \ce{^{1}H}, \ce{^{16}O}, \ce{^{56}Fe}, \ce{^{235,238}U} and \ce{^{239}Pu} in order to address the issue of compensating errors.
 Recently, this list has been expanded by the IAEA, via CIELO's follow-up, the International Nuclear Data Evaluation Network (INDEN), to include \ce{^{14,15}N}, \ce{^{9}Be}, \ce{^{23}Na}, \ce{^{59}Co}, \ce{^{58}Ni}, \ce{^{238-242}Pu}.

CIELO has relatively little energy or angle differential data to draw upon to aid in its evaluation of (n,n') on the \enquote{Big 3} Actinides, leading to \enquote{significant differences amongst the evaluations for \ce{^{239}Pu} in the fast energy range, with ENDF/B-VII.1 and JENDL-4.0 lying significantly above the JEFF-3.1 evaluation.} \cite{Cha14}.
 The \ce{^{239}Pu} data referenced in their paper are all from indirect measurements made prior to 1970 \cite{Batchelor1969, Andreev1961}.
 The case is similar for \ce{^{235}U} where it is noted that differences for $E_n >$ 50 keV \enquote{have a large impact on the fast criticality (Godiva), leading to a 540 pcm swing in calculated criticality, as shown by Go Chiba \cite{Chi12}.}

>>>>

@@

%Authors, affiliations address.
\author{Lee A. Bernstein$^{1,2}$, David A. Brown$^3$, \\Arjan J. Koning$^4$, Brad T. Rearden$^5$,  \\Catherine E. Romano$^6$, Alejandro A. Sonzogni$^3$, Andrew S. Voyles$^{1,2}$, and Walid  Younes$^7$
\affil{$^1$Nuclear Science Division, Lawrence Berkeley National Laboratory,  Berkeley, CA 94720, USA; email: labernstein@lbl.gov}
\affil{$^2$Department of Nuclear Engineering, University of California, Berkeley,  Berkeley, CA 94720, USA}
\affil{$^3$National Nuclear Data Center, Brookhaven National Laboratory, Upton, NY 11973, USA}
\affil{$^4$Arjan's affiliations\textred{Arjan - please provide!}}
\affil{$^5$Brad's affiliations\textred{Brad - please provide!}, Oak Ridge National Laboratory, Oak Ridge, TN 37830, USA}
\affil{$^6$Reactor and Nuclear Systems Division, Oak Ridge National Laboratory, Oak Ridge, TN 37830, USA}
\affil{$^7$\textred{Add Walid's division here}, Lawrence Livermore National Laboratory,  Livermore, CA 94550, USA}}

%Abstract
\begin{abstract}
A well-established knowledge of nuclear phenomena including fission, reaction cross sections, and structure/decay properties, is critical for applications ranging from the design of new reactors, to non-proliferation, to the production of radioisotopes for the diagnosis and treatment of illness.  However, the lack of a well-quantified, predictive theoretical capability means that most nuclear observables must be measured directly in order to \enquote{calibrate} empirical models, which, in turn, provide the data needed for these applications.  In many cases there is either a lack of the data needed to guide the models, or the results of the different measurements are discrepant.  This has led to the development of evaluation methodologies to provide recommended values and uncertainties.  In this article we present the nuclear data evaluation process and the international community that carries it out.  We then discuss new measurements and improved theory and/or modeling needed to address future challenges in applied nuclear science. 
\end{abstract}

@@

Low-energy nuclear science (LENS) \enquote{straddles the fence} between curiosity- and application-driven pursuits.  On the curiosity-driven side, low-energy ($E_x <$ 1--4 MeV) nuclear structure offers a unique laboratory to examine the interplay between single-particle and collective behavior and to explore the transition from quantum to continuum behavior in a mesoscopic setting. LENS also plays an important role in nuclear astrophysics, improving models of stellar energy generation and allowing isotopic abundances to be used to inform models of the astrophysical environments where heavy nuclei are formed.  On the applied side, a well-quantified knowledge of low-lying nuclear structure and reactions is needed to model energy generation and isotope production for medical and industrial uses and for the national security and counter-proliferation communities.

One hallmark of LENS is that most theoretical descriptions are descriptive rather than rigorously predictive at the level of accuracy and over the entire range of energies etc. needed for real world applications.  As a result, targeted measurements coupled to a data evaluation process are used to produce databases of recommended values that can be employed by the user communities.  The evaluation processes range from close adherence to experimental results in the case of low energy nuclear structure and decay data, to experiment-guided modeling in the case of nuclear reactions. This work is carried out by subject matter experts who usually come from the field utilizing the data who spend many years studying under experienced evaluators in order to develop the highly-specialized skills they need.

This manuscript will start with a description of the nuclear data evaluation process, and the organizations organization that carry it out.  It will then describe ways to improve the modeling of (n,x) reactions on nuclides important for applications.  Following this will be a discussion of needs for specific applications including national security/non-proliferation, medical isotope production and nuclear energy.  It will end with a discussion of the future of nuclear data featuring  a new, Nuclear Data Interagency Working Group that that is formulating a national nuclear data plan to address high-priority needs for applied nuclear science and technology.

\subsection{The Nuclear Data Pipeline}

Before diving into the data needs, it is instructive to understand the \enquote{Nuclear Data Pipeline} that takes experimental or theoretical results and prepares them for use in applications.  This process is illustrated in \autoref{fig:pipeline} and is roughly grouped into four steps: 1) compilation, 2) evaluation, 3) processing, and 4) validation.  This process ends with the users and their application, but actually begins with the performance of carefully-designed measurements with well-defined uncertainties.  Given this, it behooves those funding or performing the original experiment to understand the entire nuclear data workflow in order to ensure that new results find their way to the intended user.  For example, if a new result forces changes to data formats (step 2) or processing (step 3) or to application codes (step 4 \& eventual user), this fact should be considered when developing a new activity.

@@

The nuclear structure compilation database is called the Experimental Unevaluated Nuclear Data List (XUNDL) \cite{XUN}.   The nuclear data in XUNDL is organized by nuclide, meaning that a single publication could lead to the production of more than one XUNDL dataset.  Each XUNDL dataset is treated as a \enquote{stand-alone} work and is not required to agree with existing, evaluated nuclear data for the nuclide it concerns.  However, during XUNDL compilation basic consistency checks are performed and any internal inconsistencies are pointed out to the author to allow them be reconciled in the database.  XUNDL is used by both nuclear structure evaluators to aid in their work and by nuclear structure researchers as a quick way to get access to data from current publications in order to guide their own research.

@@

The next step in the process is evaluation itself.  In the case of nuclear structure data this involves reconciling multiple types data sets (e.g., prompt gamma and particle spectra, decay data etc.) and deriving a recommended set of adopted values.  This data forms the basis of the Evaluated Nuclear Structure Data File (ENSDF) \cite{tepel1984ensdf}.  While most of the ENSDF evaluation is performed in the US, there is a significant and growing component coming from evaluators throughout the world.  The governing body that determines the rules for the evaluation process is the Nuclear Structure and Decay Data Network administered by the International Atomic Energy Agency (IAEA-NSDD).

In the case of nuclear reaction data, the process is markedly different.  The data from EXFOR is used to guide a physics-based model calculation through a variety of variance minimization procedures, resulting in best estimates of mean values and their uncertainties, including covariances, which are put into the Evaluated Nuclear Data File \cite{Chadwick2011}.  Most of the effort put into the production of ENDF is provided by applications-oriented users due to its importance to national security, international counter-proliferation and nuclear energy.  A consequence of this application focus is that the vast majority of the data in ENDF concerns neutron-induced reactions at either thermal (25 meV), fast fission or 14.1 MeV.  This has profound implications for isotope production since many of the nuclides are produced through charged-particle induced reactions.

While ENSDF and ENDF contain a vast quantity of nuclear data, there are significant amounts of nuclear data that are either not contained in them because they \enquote{fall between} nuclear structure and nuclear reactions, or they are present in a format that is not well-suited to their use in applications.  Examples of this include the recently updated Atlas of Neutron Resonances \cite{Mug18} which contains neutron capture resonance widths, centroids and cross sections and the Evaluated Gamma Activation File of (EGAF) capture gamma-rays \cite{Fir15} and the Atlas of Inelastic Scattering of Reactor Fast Neutron \cite{Hur18}.  Another very important nuclear data resource for applications is the Reference Input Parameter Library (RIPL) \cite{Capote2009} which contains both discrete (distilled from ENSDF) and continuous (Nuclear Level Densities and Gamma-strength function) nuclear structure data needed for accurate reaction modeling.

Evaluations in hand, we must now process the data into a form suitable for use in a particular application.  This is a surprisingly non-trivial part of the process, requiring a deep knowledge of the physics used in the evaluation process, the formatting limitations of intermediate data formats, and the capabilities of downstream codes that will use the nuclear data.  Processing is also needed for the validation of nuclear reaction data using the integral benchmarks described below.  Another role of processing is in the creation of application-specific databases, such as the Medical Internal Radiation Dose (MIRD) \cite{MIRD} database used to guide medical treatment and diagnosis.

The final step in the Pipeline is validation.  In this step, we use user codes and user-defined problems to validate that the data performs as expected.  The class of user problems we most rely on are so-called \enquote{integral benchmark} data that depend on many different types of data including cross sections and emitted particle spectra and are also extremely precisely known.  The canonical example of such a benchmark is a critical assembly, where the ratio of neutron production to neutron loss ($k_{eff}$) is known to several significant figures.  However, there are many other kinds of integral benchmarks used to test neutron interaction data alone, including reaction rate measurements and 14 MeV and fission spectrum transmission experiments, and many more for other forms of nuclear data.

@@

Much of the work of nuclear data evaluation takes place as a part of international collaborations covered under the Organization of Economic Cooperation and Development's Nuclear Energy Agency (NEA) and the International Atomic Energy Agency (IAEA).  The Nuclear Energy Agency (NEA) addresses the nuclear technology interests of its 33 member states and, in the area of nuclear data, the Working Party on Evaluation Coordination (WPEC) is the main forum for collaborative effort between the nuclear data library projects from the NEA countries, namely ENDF/B (United States), JENDL (Japan), JEFF (NEA), TENDL (Europe), BROND (Russia) as well as the non-OECD file project CENDL (China).

A powerful instrument of WPEC to drive progress in nuclear data is the so-called Subgroup (SG): members of WPEC identify a common area in nuclear data that requires improvement, and if enough support from the various data library projects is present, a Subgroup is formed.  Subgroups typically operate on a 3-5 year time frame. Recent successful examples of WPEC subgroups, with specific relevance to the US, are large-scale horizontal efforts like CIELO (SG40), on worldwide nuclear data evaluation for the most important fission energy related materials. The CIELO initiative has, for example, led to new ENDF/B-VIII.0 evaluations for \ce{^{235,238}U}, \ce{^{56}Fe} and \ce{^{16}O} among others, and the Generalized Nuclear Database Structure (GNDS) format (SG38) \cite{SG38}, or more specific efforts like covariance adjustment for improvement of nuclear data files (SG39) \cite{SG39}.

WPEC also hosts long-term Expert Groups, and current groups are working on the development of the Generalized Nuclear Database Structure (GNDS) format, to provide a data library interface between nuclear physics and applications more modern than the ENDF-6 format, and the High-Priority Request List which assembles the most important nuclear data requests from applications in a unified format, to stimulate experimentalists and evaluators to provide these data. A full list of past and current subgroups of WPEC is available at Ref. \cite{WPEC}.

The International Atomic Energy Agency (IAEA) covers the interests of its 170 member states. The main task of the IAEA Nuclear Data Section is to provide fundamental nuclear data bases for basic and applied use, with data originating from experiments and theoretical simulations and covering both nuclear structure and nuclear reaction data.  An important collaboration coordinated by the IAEA is the Nuclear Reaction Data Center Network, which is responsible for keeping the EXFOR database of experimental nuclear reaction data up to date. NNDC is responsible for the US input to the database.

In addition, the IAEA organizes Coordinated Research Projects (CRP) and technical meetings as instruments to align international nuclear data efforts towards the production of validated databases ready for applied use.  Examples of recent and current CRP's include a 2018 venture centered on nuclear data for primary radiation damage has been completed, an ongoing effort to improve nuclear model parameters for fission reaction calculations by modern nuclear model codes such as EMPIRE \cite{Herman2007}, CCONE \cite{Iwa16}, COH3 \cite{KAWANO2010} and TALYS \cite{Koning2012} is ongoing, and an effort to create a first-ever evaluated database of radiative strength.  In 2019, a CRP on fission yields will start which aims to produce updated fission yield libraries for the major actinides to respond to requests from reactor technology, safeguards and non-proliferation.

@@

For fission theory, we can make a distinction in practice between description of the fission process up to scission (pre-scission physics), and the events that follow scission (post-scission physics). Eventually, one could envision a comprehensive model that accurately follows the entire fission process from the formation of the parent nucleus to the last decays of the final product nuclei within a single internally consistent theoretical framework. Such a model is not yet feasible so different physics models are used to treat pre- and post-scission processes. A wide range of theoretical approaches have been developed since the late 1930's to describe pre-scission physics, including scission-point models \cite{Wil76,Lem15}, microscopic-macroscopic models \cite{Mol15} with random-walk dynamics \cite{War17}, Langevin dynamics \cite{Sie17}, Time-dependent Hartree-Fock and Hartree-Fock-Bogoliubov theory \cite{Koo76,Neg78,Tan15,God16}, the time-dependent superfluid local density approximation \cite{Ste11,Bul13,Ste15,Bul16}, and the time-dependent generator-coordinate method \cite{Ber91,Gou05,Rei16}.Several models have also been developed to describe post-scission physics and built into computer codes, such as GEF \cite{Schm16}, FIFRELIN \cite{Lit15}, CGMF \cite{Tal14}, and FREYA \cite{Ran09,Ver18}.

The division between pre- and post-scission physics in modeling efforts suggests a corresponding division in nuclear data needs to improve our understanding of the fission process. Post-scission observables are generally easier to measure, and include the properties of fission fragments and products and of the neutrons and gamma rays they emit.  Recently, Bertsch et al. \cite{Ber15} have suggested a subset of fission data that tend to have relatively small uncertainties and could therefore be used to test and validate fission models. These data consist of fission barriers, fragment mass distributions, fragment total kinetic energies, spontaneous-fission lifetimes, and fission isomer excitation energies. In addition to these quantities, the angular distribution of fission fragments can also provide useful information about the angular momentum of the fissioning nucleus and further constrain model calculations.

@@

Both pre-scission and post-scission theory and data will be needed to reach a satisfactory understanding of the fission process, and thereby make reliable predictions of nuclear quantities that cannot be measured in the laboratory. Pre-scission models can provide input to post-scission models wherever the data are lacking, and post-scission models can be used to improve pre-scission calculations using a wealth of available experimental data. The ultimate goal for theorists, a comprehensive description of fission starting from protons, neutrons, and their interactions that provides accurate values for all fission data of interest, remains elusive.  Many questions about the fission process remain open, such as: how the initial partition of energy between kinetic and internal excitation occurs, what the relevant collective and single-particle degrees of freedom are near scission, and how they evolve and couple throughout the process? Lastly, how can we best quantify uncertainties (both epistemic and aleatory) in fission calculations?

@@

Elastic and inelastic scattering have a profound effect on neutron populations in any nuclear application by changing the directions and energies of the neutrons in the system.  However, improvements in neutron scattering data have been hampered by the fact that they do not produce easily measured signals like fission, or a radiochemical observable like neutron capture.  The result has been significant disagreements between different reaction evaluations.  These disagreements manifest themselves as compensating errors that are resistant to the validation process.  This issue has been recognized by the nuclear data community since the early days of the WPEC \cite{Row89} and more recently by Maslov \cite{Mas11}.

One compelling example of compensating errors involves the Jezebel \ce{^{239}Pu} critical assembly (PMF1).  Models of Jezebel using LANL CIELO-1 (ENDF/B-VIII.0) and CEA CIELO-2 (JEFF-3.3) reproduce a unity value of $k_{eff}$ to better than 1 part in 104.  However, when the authors substituted one nuclear quantity one quantity at a time between the two databases, the value of $k_{eff}$ varied considerably.  The greatest changes came from changing the elastic and inelastic scattering cross sections, which led to changes in the $k_{eff}$ of -500 pcm and +704 pcm, respectively. This is shown in \autoref{fig:rotated_swings}.   Similar examples of such compensating errors were also seen in earlier work by Bauge \cite{Bau12}.

\begin{figure}
 \centering
%  \includegraphics[clip=true,trim=1.5in 3.4in 1.8in 3.5in, scale=0.8]{./figures/variation_curve.pdf}
 \includegraphics[width=0.8\linewidth]{rotated_swings.pdf}

@@

A new Collaborative International Evaluated Library Organization (CIELO) Coordinated Research Program was started under the auspices of the IAEA \cite{Cha18a} to have the international nuclear data evaluation community agree on a set of evaluated cross sections and output spectra for six critically-important nuclei: \ce{^{1}H}, \ce{^{16}O}, \ce{^{56}Fe}, \ce{^{235,238}U} and \ce{^{239}Pu} in order to address the issue of compensating errors.  Recently, this list has been expanded to include \ce{^{14,15}N}, \ce{^{9}Be}, \ce{^{23}Na}, \ce{^{59}Co}, \ce{^{58}Ni}, \ce{^{238-242}Pu}.

CIELO has relatively little energy or angle differential data to draw upon to aid in its evaluation of (n,n') on the \enquote{Big 3} Actinides, leading to \enquote{significant differences amongst the evaluations for \ce{^{239}Pu} in the fast energy range, with ENDF/B-VII.1 and JENDL-4.0 lying significantly above the JEFF-3.1 evaluation.} \cite{Cha14}.  The \ce{^{239}Pu} data referenced in their paper are all from indirect measurements made prior to 1970 \cite{Batchelor1969, Andreev1961}.  The case is similar for \ce{^{235}U} where they note that differences for $E_n >$ 50 keV \enquote{have a large impact on the fast criticality (Godiva), leading to a 540 pcm swing in calculated criticality, as shown by Go Chiba \cite{Chi12}.}

One significant source of (n,xn$\gamma$) partial cross section data from 1997 to 2003 was the GEANIE collaboration (Germanium Array for Neutron-Induced Excitations) at the Los Alamos Neutron Science Center/Weapons Neutron Research (LANSCE/WNR) facility. While the primary goal of the GEANIE collaboration was the determination of the \ce{^{239}Pu}(n,2n) cross section \cite{Ber02}, it resulted in the publication of more than two dozen papers over the course of more than 15 years.  Experiments were run on all three of the major Actinides: \ce{^{235}U} \cite{You01}, \ce{^{239}Pu} \cite{Ber02} and \ce{^{238}U} \cite{Fot04}.  This includes one of the first results for $^{239}$Pu(n,n'$\gamma$), shown in \autoref{fig:239pu_nn} \cite{Ber00}.  It should be noted that GEANIE did not include any neutron detectors, producing only angle-integrated partial $\gamma$-ray cross sections.

\begin{figure}
 \centering
%  \includegraphics[clip=true,trim=1.5in 3.4in 1.8in 3.5in, scale=0.8]{./figures/variation_curve.pdf}
 \includegraphics[width=0.6\linewidth]{239pu_nn.pdf}

 % variation_curve.pdf: 612x792 pixel, 72dpi, 21.59x27.94 cm, bb=0 0 612 792
 \caption{A selection of the $^{239}$Pu(n,n' $\gamma$) partial $\gamma$-ray cross sections from [Ber00] for the two transitions de-exciting the 5/2$_{2}^{+}$ level at $E_x=$285.46 keV.  }
 \label{fig:239pu_nn}
\end{figure}
